\begin{center}
	\textbf{\Large Acknowledgment}
\end{center}

\vspace{2em}

The researchers would like to express their heartfelt gratitude to the individuals, institutions, and organizations who made the completion of this Special Problem possible:

To their adviser, Dr. Francis Dimzon, for his expert guidance, valuable insights, for teaching Computer Vision topics, and unwavering support throughout the research process.

To Prof. Jumar Cadondon, for lending his time and expertise during the early stages of this Special Problem, especially in providing assistance with his drone.

To Sir Cris Beleber, for his unwavering support from the initial conceptualization of this project to out consultations with the DPWH. His assistance was instrumental in the successful completion of this study.

To the Komsai faculty and staff, for providing a nurturing and intellectually stimulating environment.

To the university personnel who ensured the researchers' safety throughout the data collection process.

To the Department of Science and Technology (DOST), for their support and promotion of scientific research and innovation.

To the Department of Public Works and Highways (DPWH), for providing valuable data and technical insights that greatly contributed to the relevance and application of the study.

To their families and friends, for their unconditional love, patience, and encouragement.

Lastly, to the University of the Philippines Visayas, for providing the resources and environment necessary for the researchers to explore and grow.

This work stands as a testament to the collaboration, support, and trust the researchers have received. They are deeply grateful.